---
title: "506326047_stats102a_HW5"
author: "Abhijeet Rao"
date: "2024-03-15"
output: html_document
---

```{r import R file, include=FALSE}
source("506326047_stats102a_hw5.R", local = knitr::knit_global())
```

# Question 1
## Part 1
### Intervals
To find the intervals I essentially just graphed the functions and chose "nice" numbers where the roots were inside the interval of length 1. So the intervals in order are:

1. [-4,-3]
2. [3, 4]
3. [1, 2]

### Approximation of Number of Iterations:
For function 1 I believe

### Bisection Method for Each function
Below we define the bisection algorithm, I've defined it as a function as this will allow us to just run one line of code instead of repeating the same thing 3 times.
```{r defining bisection function}
bisection <- function(func, interval, max_iter){
  a = interval[1]
  b = interval[2]
  
  if (func(a) == 0){
    return (a)
  }
  if (func(b) == 0){
    return (b)
  }
  
  iter_count <- 0
  mid <- (a+b)/2
  
  while (abs(a - b) > 10^(-8)){
    mid = (a+b)/2
    if (func(a) * func(mid) < 0){
      b <- mid
    } else {
      a <- mid
    }
  }
  return (mid)
}
```

#### Function 1
```{r trying bisection}
func_1 <- function(x){
  return (x^3 + 32)
}

interval_one = c(-4,-3)

func_1(bisection(func_1, interval_one, 1000))
```

#### Function 2
```{r trying bisection}
func_2 <- function(x){
  return (x^x - 151)
}
interval_one = c(3,4)

func_2(bisection(func_2, interval_one, 1000))
```

#### Function 3

```{r trying bisection}
func_3 <- function(x){
  return (exp(-x^2) - (1/10))
}
interval_one = c(1,2)

func_3(bisection(func_3, interval_one, 1000))
```