---
title: "506326047_stats102a_HW1"
author: "Abhijeet Rao"
date: "2024-01-11"
output: html_document
---

```{r, include=FALSE}
source("506326047_stats102a_hw1.R", local = knitr::knit_global())
```

## Question 1

### A) GCD Algorithm

1.  Store `x` and `y` into a vector.
2.  Let `a` and `b` be the max and min of that vector respectively.
3.  If `b` divides `a` then return `b`.
4.  Otherwise do the following:
    a.  $$
        \begin{align*}
        x_1~\textrm{mod}~x_2 &= x_3\\
        x_2~\textrm{mod}~x_3 &= x_4\\
        x_3~\textrm{mod}~x_4 &= \cdots\\
        \vdots~~~~~\\
        x_n~\textrm{mod}~x_{n+1} &= 0
        \end{align*}
        $$
5.  Then return $x_{n+1}$ .

#### Test Cases

```{r gcd() test cases}
c(gcd(72, 8) == 8, gcd(-1, 531) == 1, gcd(47011, 73618) == 1)
```

### B) LCM Algorithm

Since we have coded in the minimum and maximum consideration into the `gcd()` function above, we do not have to worry about sorting again. The `lcm()` function hinges on these two relationships: $$
\begin{align}
  \textrm{lcm}(a, b, c) &= \textrm{lcm}(\textrm{lcm}(a,b), c)\\\\
  \textrm{lcm}(a, b) &= \frac{a\cdot b}{\textrm{gcd}(a, b)}
\end{align}
$$ This is how the algorithm works: 1. Create a while loop until the vector `x` is of length 1. 2. Store `a` and `b` as the first and second elements of `x` respectively. 3. Cut the length of `x` by 1 4. Compute and store $\textrm{lcm}(a, b) = \frac{a\cdot b}{\textrm{gcd}(a, b)}$ as the first element of `x`. 5. Repeat steps 2-4 until `x` is of size 1. 6. Return `x`.

#### Test Cases

```{r lcm() test cases}
c(lcm(c(-12, 21)) == 84, lcm(c(4789, 6123, 199)) == 5835286353)
```

## Question 2

### A) Helper Function `is_prime()` Algorithm.

1.  Create a logical vector the size of `x`.
2.  Special case for 1 and 2 and set them to `TRUE`
3.  Create a vector for each element $x_i$ of `x` call it $\vec{x_i} = [2, 3, ..., x_i - 1]$
4.  mod each element of $\vec{x_i}$ to get $x_i$ $\vec{x}_{i~\textrm{mod}} = [x_i~ \textrm{mod}~2, x_i~\textrm{mod}~3, ..., x_i~\textrm{mod}~(x_i - 1)]$
5.  If any value of $\vec{x}_{i~\textrm{mod}}$ is 0 then $x_i$ is not prime since it is divisible by something other than 1 and $x_{i}$.

#### Test Cases

```{r is_prime test cases}
c(is_prime(c(9, 7)), is_prime(c(31, 24)))
```

### B) Prime Factorization.

To create this function I split it up into two separate chunks of functions. I followed the guidance given in discussions to create

1. `exp_count()`
2. `prime_divisors()`

#### I) `exp_count()` Algorithm

1.  Take in `x` and `prime`.
2.  Start a counter called `exp_count` at 0.
3.  Start dividing `x` by `prime` and divide until there is a remainder. Advance the counter every division `prime`.
4.  Return how many times we divided by `prime`.

#### II) `prime_divisors()` Algorithm

1.  Take in `x` an integer.
2.  For all the numbers 1 through x test if it is prime.
3.  Divide `x` with the primes to see which primes divide `x`.

#### III) Final `get_factors()` Algorithm
Now all we need to do is combine the two functions together.

1. Take in `x`.
2. Get the prime divisors of `x` using the `prime_divisors()` function.
3. Store the prime divisors as a list in `primes`.
4. Get the exponents of each prime divisor using the `exp_count()` function.
5. Store the exponents of each divisor in a list `exponents`
6. Return the complete list `factors` that contains both `primes` and `exponents.`

#### Test Cases
```{r prime factorization test cases, error=TRUE}
x <- sample(x = 1e4, size = 1)
y <- get_factors(x)
this_works <- prod(y$primes^y$exponents) == x & all(is_prime(y$primes))

this_works
get_factors(1920)
get_factors(1.92)
```
