---
title: "506326047_stats102a_HW2"
author: "Abhijeet Rao"
date: "2024-01-30"
output: html_document
---

# Question 1.
```{r libraries, include=FALSE}
library(tidyverse)
source("506326047_stats102a_hw2.R", local = knitr::knit_global())
```

## Creating the Simulated Data.
Below is the code block that shows how I created the simulated data. First I used the `runif()` function to produce 1200 random values. 1200 because we have 100 rows of 5 quizzes and 7 quizzes so $100\cdot(5 + 7) = 1200$. Then we reshape that into a matrix with 12 columns. Finally we combine the matrices with sequentially generated `UIDs`.
```{r simulated data}
set.seed(506326047) ## set the seed to UID

## Create a character vector containing the names of homeworks and quizzes
names <- paste0("Homework_", 1:5)
names <- append(names, paste0("Quiz_", 1:7))

## Create the gradebook_matrix with 1200 observations and reshape it to a matrix.
gradebook_matrix <- matrix(runif(n = 1200, min = 0, max = 100), ncol = length(names), byrow = TRUE)

## Rename the columns
colnames(gradebook_matrix) <- names

## Create the gradebook dataframe with the UID column and the gradebook_matrix.
gradebook <- data.frame(UID = 123456789 + (1:100), gradebook_matrix)
```


## Creating Missing Values.
To generate the missing values we use the `sample()` function to give us 12 indices of rows between 0 and 100 and then change those values to `NA`.
```{r generating missing values homework_4}
## create the list of random rows we will replace with NA
for (i in sample(100, 12)) {
  gradebook$Homework_4[i] <- NA
}
sum(is.na(gradebook$Homework_4))
```

```{r generating missing values quiz_4}
for (i in sample(100, 12)) {
  gradebook$Quiz_4[i] <- NA
}
sum(is.na(gradebook$Quiz_4))
```

```{r tibble conversion}
gradebook <- as_tibble(gradebook)
```

## The `messy_impute()` function.
### Algorithm
I will speak about the algorithm from the `center = "mean"` initially since the `center = "median"` is essentially the exact same, just use the `median()` function instead. Throughout all these calculations for means we are ignoring `NA` values by using the `na.rm = TRUE` argument.

The key part of this function's algorithm is the following line:
```{r key line, eval=FALSE}
na_matrix <- which(is.na(input_tibble), arr.ind = TRUE)
```

This line allows stores the indices of the `NA` values in a matrix with the row and column index of the `NA` value.

#### Row-By-Row
Here is the algorithm for the row-by-row/student `NA` impute;

1. If `margin = 1` then do the row-by-row replacement of `NA` values which is the following;
2. For each row in the `na_matrix` first check if the row is a Homework assignment or Quiz Assignment just by seeing if the column of the `NA` value we are checking is between columns 2-6  or 7-13.
3. To make the code more readable, store the row as a tibble into the `placeholder_row` variable.
4. Remove the names on the row so we can actually perform the mean function using the `unlist(..., use.names=FALSE)`
5. If we are replacing a Homework then use only the homework row by slicing indices 2-6 of the row. If we are replacing a Quiz then use the quiz row by slicing indices 7-13 of the row.
5. Calculate the mean of the row and input it into the cell that we were looking at with the `NA` value.
6. Repeat this for each `NA` value.
7. Return the tibble.

#### Column
Next, here is the algorithm for the column/assignment `NA` impute:

side-note: this one was much simpler as column operations are very easy to do with tibbles whereas rows are a bit tougher to work with.

1. If `margin = 2` then do the column replacement of `NA` values which is the following;
2. For each column in the `na_matrix` take the mean of the entire column and input it into the missing value cell.
3. Repeat for each `NA` value.
4. Return the tibble

#### Test Case
Here I showcase the student with UID 123456797 and the imputation of their `NA` scores for the Homework 4 assignments.

```{r test-case}
gradebook[8,]
messy_impute(gradebook, "mean", 1)[8, ]
messy_impute(gradebook, "median", 2)[8, ]
messy_impute(gradebook, "mean", 1, trim = 0.25)[8, ]
```

## Conversion into tidy data.
To convert the `gradebook` data set into tidy data we use the `pivot_longer()` function. Additionally make sure we didn't lose any `NA` values through our test cases before hand.

```{r tidy time}
gradebook_tidy <- gradebook %>% pivot_longer(cols = !UID,
                           names_to = "Assignments",
                           values_to = "Marks")
gradebook_tidy
```

```{r trying out tidy_impute}
tidy_impute(gradebook_tidy, center = "mean", margin = 1)
```



<!-- input tibble
which(is.na(tibble), arr.ind = TRUE)
returns a matrix of two columns 
first column will give row numbers and second column will give column numbers
what are the row and column indices of each of those trues.
go row by row and find the positions of each of those in the data frame.


sends column names to a new variable and stores the price into a different column
grocery_tbl %>% pivot_longer(cols = !fruit(all columns except for fruit),
                             names_to = "supermarket",
                             values_to = "price")
                             
                             
The tidy object that you get out should have three columns
UID, Assignments, Marks.


check out harvard dataverse for q2
-->
