---
title: "506326047_stats102a_HW2"
author: "Abhijeet Rao"
date: "2024-01-30"
output: html_document
---

# Question 1.
```{r libraries, include=FALSE}
library(tidyverse)
source("506326047_stats102a_hw2.R", local = knitr::knit_global())
```

## Creating the Simulated Data.
Below is the code block that shows how I created the simulated data. First I used the `runif()` function to produce 1200 random values. 1200 because we have 100 rows of 5 quizzes and 7 quizzes so $100\cdot(5 + 7) = 1200$. Then we reshape that into a matrix with 12 columns. Finally we combine the matrices with sequentially generated `UIDs`.
```{r simulated data}
set.seed(506326047) ## set the seed to UID

## Create a character vector containing the names of homeworks and quizzes
names <- paste0("Homework_", 1:5)
names <- append(names, paste0("Quiz_", 1:7))

## Create the gradebook_matrix with 1200 observations and reshape it to a matrix.
gradebook_matrix <- matrix(runif(n=1200, min = 0, max = 100), ncol=length(names), byrow=TRUE)

## Rename the columns
colnames(gradebook_matrix) <- names

## Create the gradebook dataframe with the UID column and the gradebook_matrix.
gradebook <- data.frame(UID=123456789+(1:100), gradebook_matrix)
```

## Creating Missing Values.
To generate the missing values we use the `sample()` function to give us 12 indices of rows between 0 and 100 and then change those values to `NA`.
```{r generating missing values homework_4}
## create the list of random rows we will replace with NA
for (i in sample(100, 12)){
  gradebook$Homework_4[i] = NA
}
sum(is.na(gradebook$Homework_4))
```

```{r generating missing values quiz_4}
for (i in sample(100, 12)){
  gradebook$Quiz_4[i] = NA
}
sum(is.na(gradebook$Quiz_4))
```

## The `messy_impute()` function.
### Algorithm
1. Firstly we have some error handling steps

input tibble
which(is.na(tibble), arr.ind = TRUE)
returns a matrix of two columns 
first column will give row numbers and second column will give column numbers
what are the row and column indices of each of those trues.
go row by row and find the positions of each of those in the data frame.


sends column names to a new variable and stores the price into a different column
grocery_tbl %>% pivot_longer(cols = !fruit(all columns except for fruit),
                             names_to = "supermarket",
                             values_to = "price")
                             
                             
The tidy object that you get out should have three columns
UID, Assignments, Marks.




check out harvard dataverse for q2