---
title: "506326047_stats102a_HW4"
author: "Abhijeet Rao"
date: "2024-03-01"
output: html_document
---

```{r import R file, include=FALSE}
source("506326047_stats102a_hw4.R", local = knitr::knit_global())
```

## Algorithm for Required Functions.
### Carry-Over Function.
1. Take in two pq numbers
2. Set `p` and `q` to be the max `p` and `q` of the two `pqnumber`s
3. Store the `nums`, `p`, `q` of both `pqnumber`s
4. Pad the `pqnumber`s appropriately.
    a. check if the `p`(`q`) of a `pqnumber` is less than the max `p`(`q`)
    b. if it is less, then add a 0 at the end(beginning) and continue until they are equal.
5. Now that the two `nums` vectors have the proper dimensions, add them together.
6. Loop through each element in the summed vector and if its more than 9, subtract 10 and add one to the next element in the list.
7. Finally if the last element is greater than 9, then add a final 1 at the end of the vector.
8. Return the `p` and `q` that were calculated along with the summed vector as a named list.

### Addition Function.
1. Take in two `pqnumber`s.
2. If the signs are equal then it is an addition, otherwise its a subtraction.
3. Store the sign of the `pqnumber`s and re-assign them to 1 so we can use the `carry_over` function.
4. Use the results from the `carry_over` and the stored sign to create a new `pqnumber` object.

### Borrowing Function.
This function follows a similar principle as the `carry_over` function.

NOTE:  In this function it is assumed that the first `pqnumber` is greater than the second `pqnumber`

1. Assign `p` and `q` values as the max `p` and `q` of the two `pqnumber`s
2. Pad them just as we did for the `carry_over` function.
3. Subtract the two `nums`.
4. If an element in the subtracted list is less than 0 then add 10 and subtract 1 from the next element. Do this up till the last element.
5. Return the `p`, `q` and the subtractedv ector as a named list.

### Subtraction Function.
1. In subtraction, if the signs of two numbers are different, then it is an addition.
    a. Store the sign of the first `pqnumber` and re-assign both `pqnumber`s to have `sign = 1`
2. If the two signs are equal then pass to the next stage.
3. Test which number is bigger, then pass that number into the `borrow` function.
4. Use the returned values from the `borrow` function to create a `pqnumber` using the appropriate sign.

### Multiplication Function.

I wasn't able to figure out how to do this function without using the coercion functions. I think this may be the appropriate steps:

1. Seperate the two `pqnumber`s by their base 10s.
2. Using a for loop multiply over each base 10 throughout the vector.
3. Add each of them together.
4. Use strsplit and string to find the period and create our p,q based off this location.
5. create a new `pqnumber` based off these values.

## Test Cases.

```{r instantiating pq_numbers}
pq_1 = pqnumber(1, 3, 4, 1:8)
pq_2 = pqnumber(1, 6, 0, nums = c(3, 9, 5, 1, 4, 1, 3))
pq_3 = pqnumber(-1, 5, 1, nums = c(2, 8, 2, 8, 1, 7, 2))
```

### Add Function.
```{r printing pqnumbers}
c(print(pq_1, DEC=TRUE), print(pq_2, DEC=TRUE), print(pq_3, DEC=TRUE))
```

```{r testing add}
# adding pq_1 and pq_2
c(print(pq_1, DEC=TRUE) + print(pq_2, DEC=TRUE), print(add(pq_1, pq_2), DEC=TRUE))
# adding pq_1 and pq_3
c(print(pq_1, DEC=TRUE) + print(pq_3, DEC=TRUE), print(add(pq_1, pq_3), DEC=TRUE))
# adding pq_2 and pq_3
c(print(pq_2, DEC=TRUE) + print(pq_3, DEC=TRUE), print(add(pq_2, pq_3), DEC=TRUE))
```
### Subtraction Function.
```{r printing pqnumbers again}
c(print(pq_1, DEC=TRUE), print(pq_2, DEC=TRUE), print(pq_3, DEC=TRUE))
```

```{r testing subtract}
# subtracting pq_1 and pq_2
c(print(pq_1, DEC=TRUE) - print(pq_2, DEC=TRUE), print(subtract(pq_1, pq_2), DEC=TRUE))
# subtracting pq_1 and pq_3
c(print(pq_1, DEC=TRUE) - print(pq_3, DEC=TRUE), print(subtract(pq_1, pq_3), DEC=TRUE))
# subtracting pq_2 and pq_3
c(print(pq_2, DEC=TRUE) - print(pq_3, DEC=TRUE), print(subtract(pq_2, pq_3), DEC=TRUE))
```
